### 2020-11-10 latex/acrobot.tex
Finished the first writeup of rotations, added another section for full energy gain

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
enough \(I\), the Poincar\'{e} section \(P(\mu_0)\) is positive, monotonically
increasing, and \(\lim \limits_{\mu_0 \to \pi} P(\mu_0) = \infty\).
Thus, there are no closed orbits and the origin must be a repeller.
Furthermore, for any non-zero initial condition of the acrobot and for small
enough \(I\), the \((q_u,p_u)\) phase of the acrobot will eventually exit \(D\)
and the acrobot will begin rotating.
The acrobot's orbit must leave all compact sets in \(D\) in finite time, which
means the acrobot is gaining energy on \(D\).

@@

Take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) and initial
\(\mu_0\).
We expand the solution of this into the same form as
(\ref{eqn:acrobot-muhat-approx}) where once again we have
\(\hat{\mu}_0(\alpha) \equiv \mu_0\).
The dynamics for \(\hat{\mu}_1\) are
\begin{equation}\label{eqn:acrobot-rot-mu1-dot}
    \dot{\hat{\mu}}_1 =
    \frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
        - c_\alpha\mu_0^2
    \right)}{
    |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
    }
    .
\end{equation}
This is an even, \(2\pi\)-periodic function in \(\alpha\).

>>>>

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for small
enough \(I\), the Poincar\'{e} section \(P(\mu_0)\) is positive, monotonically
increasing, and \(\lim \limits_{\mu_0 \to \pi} P(\mu_0) = \infty\).
Thus, there are no closed orbits and the origin must be a repeller.
Furthermore, for any non-zero initial condition of the acrobot and for small
enough \(I\), \((q_u,p_u)\) will eventually exit \(D\).
This implies the orbit will escape any compact set in \(D\) in finite time. 
By Definition \ref{defn:energy-gain}, the acrobot is gaining energy on \(D\).

@@

Take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) and initial
state \(\mu_0\).
We expand the solution of this into the same form as
(\ref{eqn:acrobot-muhat-approx}) where once again we have
\(\hat{\mu}_0(\alpha) \equiv \mu_0\).
The dynamics for \(\hat{\mu}_1\) are
\begin{equation}\label{eqn:acrobot-rot-mu1-dot}
    \dot{\hat{\mu}}_1 =
    \frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
        - c_\alpha\mu_0^2
    \right)}{
    |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
    }
    =: b(\alpha,\mu_0)
    .
\end{equation}
The solution to (\ref{eqn:acrobot-rot-mu1-dot}) is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) = \int \limits_0^\alpha b(\sigma,\mu_0)d\sigma
    .
\]
Since one full rotation amounts to \(\alpha\) traversing \(2\pi\)rad in a clockwise
direction (\ie \(\alpha\) going from \(0\) to \(\sign{\mu_0}2\pi\)), 
we define a Poincar\'{e} section as the point reached on the \(p_u\) axis after
every rotation:
\[
    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\sign{\mu_0}2\pi, \mu_0)
    .
\]
If \(|P(\mu_0)| > |\mu_0|\), then the rotation has
driven the orbit further from the origin.
In other words, \(|p_u|\) has increased across the rotation.
If this is true for any \(\mu_0\), the acrobot is gaining energy.

The function \(b(\alpha,\mu_0)\) is even in both its inputs, and is 
\(2\pi\)-periodic in \(\alpha\).
Hence, we have 
\begin{align*}
    \hat{\mu}_1(\sign{\mu_0}2\pi, \mu_0) &=
    \int \limits_0^{\sign{\mu_0}2\pi} b(\sigma,\mu_0)d\sigma
    , \\
     &=\sign{\mu_0}\int\limits_0^{2\pi}b(\sigma,\mu_0)d\sigma
     , \\
     &= \sign{\mu_0} \hat{\mu}_1(2\pi,\mu_0)
     .
\end{align*}
Furthermore, \(\hat{\mu}_1(2\pi,\mu_0) = \hat{\mu}_1(2\pi,|\mu_0|)\).
This implies that
\[
    |P(\mu_0)| = \left|\sign{\mu_0}\left(
        |\mu_0| + I \hat{\mu}_1(2\pi,|\mu_0|)\right)
    \right|
    = P(|\mu_0|)
    .
\]
Therefore, if \(P(|\mu_0|) > |\mu_0|\), then the acrobot is gaining energy.
Showing that \(\hat{\mu}_1(2\pi,\mu_0) > 0\) for \(\mu_0 > \sqrt{60m^2gl^3}\) is
sufficient for this condition to hold.

Unlike in the case of oscillations, we cannot adimensionalize
\(b(\alpha,\mu_0)\) and prove that this condition will be true for all acrobots. 
Let \(U = ]\sqrt{60m^2gl^3},\bar{\mu}[\), and assume that 
\(\hat{\mu}_1(2\pi,\mu_0) > 0 \) for all \(\mu_0 \in U\).
If this is true, then by perturbation analysis we can claim that there exists a
small enough \(I\) such that the Poincar\'{e} section \(P(\mu_0)\) is monotonic
on \(U\).

Define the set of points in \(\Gamma\) which have pseudo-radius in \(U\) by
\[
    W := \left\{(q_u,p_u) \in \Gamma \mid |\mu(q_u,p_u)| \in U \right\}
    ,
\]
as in Figure \ref{fig:acrobot-rot-W}.
Monotonicity of \(P(\mu_0)\) means there are no closed orbits of 
\((q_u,p_u)\) on \(W\). 
Furthermore, \(|p_u|\) is increasing every time the orbit hits the
\(p_u\)-axis.

\begin{figure}
    \centering
    % TODO: Display the set W, which is the two rotation bands at the top and
    % bottom with outer limit +- \bar{mu}  and inner limit +- sqrt{60m^2gl^3}.
    % Show also that the inner limit connects to D
    %\includestandalone[width=0.5\textwidth]{images/acrobot_rot_W}
    \caption{The set \(W\) of points which have pseudo-radius 
        \(|\mu| \in ]\sqrt{60m^2gl^3}, \bar{\mu}[\). The inner boundary of \(W\)
        is \((q_u,p_u) = (\pi,0)\), which is also the boundary of the
        oscillation domain \(D\).}
    \label{fig:acrobot-rot-W}
\end{figure}

We conclude that \(|\mu(q_u,p_u)|\) will eventually reach \(\bar{\mu}\). 
We cannot claim this fits the definition  of energy injection 
(since \((0,0) \not \in W\)), but we can still say that \((q_u,p_u)\) will
eventually depart \(W\) through the outer boundary in finite time.
The acrobot is thereby gaining energy in the literal sense that its momentum
is increasing on \(W\).

If we set \(I < 0\), we have \(P(|\mu_0|) < |\mu_0|\).
Hence, \(|p_u|\) is decreasing every time the orbit hits the \(p_u\)-axis, so
the acrobot will eventually depart \(W\) through the inner boundary. 
The acrobot is losing momentum, and therefore it is losing energy on \(W\).

\subsection*{Connecting Oscillations and Rotations}
To complete this proof, we will prove the acrobot gains energy on the set 
\[
    E_\bar{\mu} := 
    \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) < E(0,\bar{\mu}) \right\}
    .
\]
% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
% every initial condition starting in \(D\) will NOT converge to (pi,0), and
% therefore exit D and enter W.
